Processing of visual and non-visual naturalistic spatial information in the "parahippocampal place area"

### 2021-12-03 paper/main.tex
Minor edits for readability, remove comments

% brain mapping via fMRI
Studies in the field of neuropsychology and neuroimaging
(\citep[e.g.,][]{penfield1950cerebral, fox1984noninvasive}) have shown that
different parts of the brain are specialized for different perceptual and
cognitive functions.
% occipital cortex for vision -> two pathways
The occipital cortex is considered to be primarily involved in the early stages
of visual perception and giving rise to two distinct, but interacting, pathways
that serve different functions:
% dorsal vs ventral pathway
a) a dorsal stream (the ``how pathway'') that leads into the parietal lobe and
supports visual guidance of action, and b) a ventral stream (the ``what
pathway'') that leads into the temporal lobe and supports conscious perception
and recognition \citep{goodale1992separate, milner2017two, ungerleider1982two}.
% PPA
A classic example of a higher-level visual area in the ventral pathway is the
``parahippocampal place area'' (PPA) \citep{epstein1998ppa,
epstein1999parahippocampal}.
% anatomical location != functional location
The PPA is located in the posterior parahippocampal gyrus including adjacent
regions of the fusiform gyrus and anterior lingual gyrus
\citep{epstein2008parahippocampal}.
% neural correlate of scene perception
Increased hemodynamic activity is observed in the PPA when participants view
pictures of landscapes, buildings or landmarks (compared to e.g. pictures of
faces or tools) during blood oxygenation level-dependent functional magnetic
resonance imaging (BOLD fMRI) (\citep[see reviews][]{epstein2014neural,
aminoff2013role}).
% \citep{aguirre1998area, epstein2014neural, epstein1998ppa, troiani2012object}.

% literature review overview: imagination & haptic exploration; van der Hurk: 64
% sounds x 1.8 seconds; 20 sighted, 14 congenitally blind
Increased hemodynamic activity in the PPA generalizes from pictures to mental
imagination of landscapes \citep{ocraven2000mental}, haptic exploration of
scenes constructed from LEGO blocks \citep{wolbers2011modality}, and
scene-related sounds \citep{van2017development}.
% % O'Craven: watching pictures
In a study conducted by \citet{ocraven2000mental} participants viewed
alternating blocks of pictures showing familiar places and famous faces during
an initial experimental paradigm.
% O'Craven: mental imagery
In a subsequent paradigm, participants were instructed to ``form a vivid mental
image'' of the previously viewed pictures.
% O'Craven results
The PPA showed increased activation during imagination of places compared to
faces but the imagination task showed a smaller activation level compared to the
perceptual task.
% Wolbers: haptic exploration
In a block design study conducted by \citet{wolbers2011modality} the PPA of
sighted as well as blind participants showed increased activation during a
delayed match-to-sample task of haptically explored scenes constructed from LEGO
bricks compared to abstract geometric objects.
% Wolbers: connectivity analysis

@@

% open question
In this study, we investigate whether increased hemodynamic activity in the PPA
that is usually detected by contrasting blocks of pictures is also present under
more natural conditions.
% we use 2 naturalistic stimuli
To answer this question, we operationalized the perception of both
\textit{visual and auditory} spatial information using two naturalistic stimuli
(\citep[see reviews][]{hamilton2018revolution, hasson2008neurocinematics,
sonkusare2019naturalistic}).
% AV stimulus
The current operationalization of visual spatial perception is based on an
annotation of cuts and depicted locations in the audio-visual movie ``Forrest
Gump'' \citep{haeusler2016cutanno}, while
% AD stimulus
the operationalization of non-visual spatial perception is based on an
annotation of speech occurring in the movie's audio-description
\citep{haeusler2021studyforrest}.
% pictures -> movie -> audio-description
The movie stimulus shares the stimulation in the visual domain with classical
localizer stimuli, while featuring real-life-like visual complexity and
naturalistic auditory stimulation. The audio-description maintains the
naturalistic nature of the movie stimulus, but limited to the auditory domain.
% current GLM contrasts
We applied model-based, mass-univariate analyses to BOLD fMRI data from both
naturalistic stimuli \citep{hanke2016simultaneous, hanke2014audiomovie},
available from the open-data resource
\href{http://www.studyforrest.org}{studyforrest.org}
\citep{hanke2016aligned}.\todo{ref. to GitHub aligned}
% comparison to previous GLM
We compare current results to results of a previously performed model-based,
mass-univariate analysis that was applied to data from a conventional functional
localizer performed with the same set of participants
\citep{sengupta2016extension, sengupta2016extensiondata}.\todo{new ref. to paper
\& GitHub}
% intro to hypo a
Similarly to the functional localizer, we currently also capitalize on events
that should evoke the cognitive processing of spatial information.
% hypo a
Thus, we hypothesized that our whole-brain analyses would reveal increased
hemodynamic activity in medial temporal regions that were functionally
identified as the PPA by the analysis of the localizer data.
%hypo b: individuals
We hypothesized further that a purely auditory stimulus could, in principle,
localize the PPA as an example of a ``visual area'' in individual persons,
% application in individuals
and may offer an alternative paradigm to assess brain functions in visually
impaired individuals.

@@

\todo[inline]{I tried to keep the summary of methods short and use a pretty
similar phrasing; still, I wouldn't get the result section and read the Method
section and tables there while reading the results\dots. Most important imo is
to link the Tables (Events + Contrasts) and explain that there is that ``primary
contrast'' thing.}

% study goal
We investigated if spatial information embedded in two naturalistic stimuli
correlates with increased hemodynamic activity in the PPA.
%
Based on an annotation of cuts in the movie and an annotation of speech spoken
by the audio-description's narrator, we selected events in both stimuli that
should correlate with the perception of spatial information and contrasted them
with events that should not (or to a less degree) correlate with non-spatial
perception (see Table~\ref{tab:events}).
% from supplementals
In order to test the robustness of our approach, we created multiple linear
model (GLM) $t$-contrasts for the movie and audio-description (see
Table~\ref{tab:contrasts}).
% subjectively assessed balance
For each stimulus, we chose a primary contrast for result presentation (marked
with an asterisk in the table) based on a subjectively assessed balance of how
well the averaged events within categories represent spatial and non-spatial
information, and the number of events in the stimulus.
% group level
On a group average level, we here report results from whole-brain analyses of
the movie's and the audio-description's primary contrasts and compare these
results to a traditional visual localizer \citep{sengupta2016extension}.
% individual level
For each individual, we also compare the activation pattern from naturalistic
auditory stimulation to the individual PPA localization with a block-design
stimulus.
% reference to supplementals
An evaluation of the results' robustness across all created contrasts on a group
average level is provided in the Supplementary Information (see
Figure~\ref{fig:stability-slices-volsurf}).
% NeuroVault
Unthresholded $Z$-maps of all contrasts on a group average level as well as
individual results of the primary $t$-contrasts co-registered to the
group-template (MNI152 space) can be found at
\href{https://neurovault.org/collections/KADGMGVZ/}{\url{neurovault.org/collections/KADGMGVZ}}.

@@

\todo[inline]{``method'' of creating the surface plot is written here in results
because it's just one sentence}

@@

%#13 natural stimulation
In summary, natural stimuli like movies \citep{eickhoff2020towards,
hasson2008neurocinematics, sonkusare2019naturalistic} or narratives
\citep{hamilton2018revolution, honey2012not, lerner2011topographic,
silbert2014coupled, wilson2008beyond} can be used as a continuous, complex,
immersive, task-free paradigm that more closely resembles our natural dynamic
environment than traditional experimental paradigms.
% method
We took advantage of three\todo{actually two} fMRI datasets and two stimulus annotations that are
part of the open-data resource
\href{http://www.studyforrest.org}{studyforrest.org} to operationalize the
perception of spatial information embedded in an audio-visual movie and an
auditory narrative, and compare current results to a previous report of a
conventional, block-design localizer.
% results
The current study offers evidence that a model-driven GLM analysis based on
annotations can be applied to a naturalistic paradigm to localize concise
functional areas and networks correlating with specific perceptual processes
-- an analysis approach that can be facilitated by the neuroscout.org platform
\citep{delavega2021neuroscout}.
% interpretation
More specifically, our results demonstrate that increased activation in the PPA
during the perception of static pictures generalizes to the perception of
spatial information embedded in a movie and an exclusively auditory stimulus.
% interpretation: aPPA vs. pPPA
Our results provide further evidence that the PPA can be divided into
functional subregions that coactivate during the perception of visual scenes.
% interpretation
Finally, the presented evidence on the in-principle suitability of a naturally
engaging, purely auditory paradigm for localizing the PPA may offer a path to
the development of diagnostic procedures more suitable for individuals with
visual impairments or conditions like nystagmus.

>>>>

% brain mapping via fMRI
Studies in the field of neuropsychology and neuroimaging
(\citep[e.g.,][]{penfield1950cerebral, fox1984noninvasive}) have shown that
different parts of the brain are specialized for different perceptual and
cognitive functions.
% occipital cortex for vision -> two pathways
The occipital cortex is considered to be primarily involved in the early stages
of visual perception and giving rise to two distinct, but interacting, pathways
that serve different functions:
% dorsal vs ventral pathway
a) a dorsal stream (the ``how pathway'') that leads into the parietal lobe and
supports visual guidance of action, and b) a ventral stream (the ``what
pathway'') that leads into the temporal lobe and supports conscious perception
and recognition \citep{goodale1992separate, milner2017two, ungerleider1982two}.
% PPA
A classic example of a higher-level visual area in the ventral pathway is the
``parahippocampal place area'' (PPA) \citep{epstein1998ppa,
epstein1999parahippocampal}.
% anatomical location != functional location
The PPA is located in the posterior parahippocampal gyrus including adjacent
regions of the fusiform gyrus and anterior lingual gyrus
\citep{epstein2008parahippocampal}.
% neural correlate of scene perception
Increased hemodynamic activity is observed in the PPA when participants view
pictures of landscapes, buildings or landmarks, compared to e.g. pictures of
faces or tools, during blood oxygenation level-dependent functional magnetic
resonance imaging (BOLD fMRI) (\citep[see reviews][]{epstein2014neural,
aminoff2013role}).
% \citep{aguirre1998area, epstein2014neural, epstein1998ppa, troiani2012object}.

% literature review overview: imagination & haptic exploration; van der Hurk: 64
% sounds x 1.8 seconds; 20 sighted, 14 congenitally blind
Increased hemodynamic activity in the PPA generalizes from pictures to mental
imagery of landscapes \citep{ocraven2000mental}, haptic exploration of
scenes constructed from LEGO blocks \citep{wolbers2011modality}, and
scene-related sounds \citep{van2017development}.
% % O'Craven: watching pictures
In a study conducted by \citet{ocraven2000mental} participants viewed
alternating blocks of pictures showing familiar places and famous faces during
an initial experimental paradigm.
% O'Craven: mental imagery
In a subsequent paradigm, participants were instructed to ``form a vivid mental
image'' of the previously viewed pictures.
% O'Craven results
The PPA showed increased activation during imagination of places compared to
faces but the imagination task showed a smaller activation level compared to the
perceptual task.
% Wolbers: haptic exploration
In a block design study conducted by \citet{wolbers2011modality} the PPA of
sighted as well as blind participants showed increased activation during a
delayed match-to-sample task of haptically explored scenes constructed from LEGO
bricks compared to abstract geometric objects.
% Wolbers: connectivity analysis

@@

% open question
In this study, we investigate whether increased hemodynamic activity in the PPA
that is usually detected by contrasting blocks of pictures is also present under
more natural conditions.
% we use 2 naturalistic stimuli
To answer this question, we operationalized the perception of both
\textit{visual and auditory} spatial information using two naturalistic stimuli
(\citep[see reviews][]{hamilton2018revolution, hasson2008neurocinematics,
sonkusare2019naturalistic}).
% AV stimulus
The current operationalization of visual spatial perception is based on an
annotation of cuts and depicted locations in the audio-visual movie ``Forrest
Gump'' \citep{haeusler2016cutanno}, while
% AD stimulus
the operationalization of non-visual spatial perception is based on an
annotation of speech occurring in the movie's audio-description
\citep{haeusler2021studyforrest}.
% pictures -> movie -> audio-description
The movie stimulus shares the stimulation in the visual domain with classical
localizer stimuli, while featuring real-life-like visual complexity and
naturalistic auditory stimulation. The audio-description maintains the
naturalistic nature of the movie stimulus, but limited to the auditory domain.
% current GLM contrasts
We applied model-based, mass-univariate analyses to BOLD fMRI data from both
naturalistic stimuli \citep{hanke2016simultaneous, hanke2014audiomovie},
available from the open-data resource
\href{http://www.studyforrest.org}{studyforrest.org}
\citep{hanke2016aligned}.\todo{ref. to GitHub aligned}
% comparison to previous GLM
We compare current results to results of a previously performed model-based,
mass-univariate analysis that was applied to data from a conventional functional
localizer performed with the same set of participants
\citep{sengupta2016extension, sengupta2016extensiondata}.\todo{new ref. to paper
\& GitHub}
% intro to hypo a
Similarly to the functional localizer, we also capitalize on events
that should evoke the cognitive processing of spatial information.
% hypo a
Thus, we hypothesized that our whole-brain analyses would reveal increased
hemodynamic activity in medial temporal regions that were functionally
identified as the PPA by the analysis of the localizer data.
%hypo b: individuals
We hypothesized further that a purely auditory stimulus could, in principle,
localize the PPA as an example of a ``visual area'' in individual persons,
% application in individuals
and may offer an alternative paradigm to assess brain functions in visually
impaired individuals.

@@

% study goal
We investigated if spatial information embedded in two naturalistic stimuli
correlates with increased hemodynamic activity in the PPA.
%
Based on an annotation of cuts in the movie and an annotation of speech spoken
by the audio-description's narrator, we selected events in both stimuli that
should correlate with the perception of spatial information and contrasted them
with events that should correlate with non-spatial
perception to a lesser degree, or not at all (see Table~\ref{tab:events}).
% from supplementals
In order to test the robustness of our approach, we created multiple linear
model (GLM) $t$-contrasts for the movie and audio-description (see
Table~\ref{tab:contrasts}).
% subjectively assessed balance
For each stimulus, we chose a primary contrast for result presentation
based on a subjectively assessed balance of how
well the averaged events within categories represent spatial and non-spatial
information, and the number of events in the stimulus.
% group level
On a group average level, we report results from whole-brain analyses of
the movie's and the audio-description's primary contrasts and compare these
results to a traditional visual localizer \citep{sengupta2016extension}.
% individual level
For each individual, we also compare the activation pattern from naturalistic
auditory stimulation to the individual PPA localization with a block-design
stimulus.
% reference to supplementals
An evaluation of the results' robustness across all created contrasts on a group
average level is provided in the Supplementary Information (see
Figure~\ref{fig:stability-slices-volsurf}).
% NeuroVault
Unthresholded $Z$-maps of all contrasts on a group average level as well as
individual results of the primary $t$-contrasts co-registered to the
group-template (MNI152 space) can be found at
\href{https://neurovault.org/collections/KADGMGVZ/}{\url{neurovault.org/collections/KADGMGVZ}}.

@@

@@

%#13 natural stimulation
In summary, natural stimuli like movies \citep{eickhoff2020towards,
hasson2008neurocinematics, sonkusare2019naturalistic} or narratives
\citep{hamilton2018revolution, honey2012not, lerner2011topographic,
silbert2014coupled, wilson2008beyond} can be used as a continuous, complex,
immersive, task-free paradigm that more closely resembles our natural dynamic
environment than traditional experimental paradigms.
% method
We took advantage of three fMRI acquisitions and two stimulus annotations that are
part of the open-data resource
\href{http://www.studyforrest.org}{studyforrest.org} to operationalize the
perception of spatial information embedded in an audio-visual movie and an
auditory narrative, and compare current results to a previous report of a
conventional, block-design localizer.
% results
The current study offers evidence that a model-driven GLM analysis based on
annotations can be applied to a naturalistic paradigm to localize concise
functional areas and networks correlating with specific perceptual processes
-- an analysis approach that can be facilitated by the neuroscout.org platform
\citep{delavega2021neuroscout}.
% interpretation
More specifically, our results demonstrate that increased activation in the PPA
during the perception of static pictures generalizes to the perception of
spatial information embedded in a movie and an exclusively auditory stimulus.
% interpretation: aPPA vs. pPPA
Our results provide further evidence that the PPA can be divided into
functional subregions that coactivate during the perception of visual scenes.
% interpretation
Finally, the presented evidence on the in-principle suitability of a naturally
engaging, purely auditory paradigm for localizing the PPA may offer a path to
the development of diagnostic procedures more suitable for individuals with
visual impairments or conditions like nystagmus.
